Concept note: Maragra model approach - September 2018

### 2020-06-20 inst/rmd/sep2018/sep2018.tex
updated

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Concept note: additional models},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

\setlength{\droptitle}{-2em}
  \title{Concept note: additional models}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
\subtitle{Pradhan, Sicuri, Brew}
  \author{}
  \preauthor{}\postauthor{}
  \date{}
  \predate{}\postdate{}

@@

A few different approaches to modeling absenteeism in the Maragra data

@@

This document contains methods and basic results for a few different modeling approaches, following a conversation and write-up with Menno Pradhan in May 2018. The purpose of this document is to make explicit the current models we have implemented, and some issues with them.

@@

\noindent\fbox{%
    \parbox{\textwidth}{%
        \subsection*{Main points}
        \begin{itemize}
          \item I have implemented Menno's suggestions for more simple models.
          \item The simplest models show no overall effect from IRS when not accounting for malaria season
          \item I believe that the lack of an effect is due to seasonality's presence in the model as a fixed (additive) effect, rather than as an interaction with IRS
        \end{itemize}
        \vspace{2mm}
    }%
}

@@

\section{The current approach}\label{the-current-approach}

In the current draft of the paper
(\href{https://docs.google.com/document/d/1bUWRBCgVcgjSPHchIQxiTG8Vwv5hV1GLU4Tlu386sWA/edit#}{HERE}
), the model formula is:

@@

\begin{verbatim}
felm(formula = absent ~ season * months_since + rainy_day + herd | 
    oracle_number + malaria_year | 0 | 0, data = these_data)
\end{verbatim}

@@

\section{Suggested changes from
Menno}\label{suggested-changes-from-menno}

Menno sent a write-up on May 17th (see email with subject ``suggestions
for models'') with some proposed changes to how we model absenteeism.
What follows is an implementation of those suggestions.

\subsection{Model A (without
seasonality)}\label{model-a-without-seasonality}

The below is a characterization of the approach Menno suggested:

\begin{itemize}
\tightlist
\item
  We use monthly instead of daily data (modeling a percentage
  absenteeism rate rather than a binary).
\item
  We use fixed effects for year-month combination.
\item
  We don't explicitly model for malaria season.
\item
  We don't have any interaction terms.
\item
  We use dummies for months since IRS.
\item
  We use the most recent spraying as the time since IRS (ie, if 5 months
  ago, and 1 month ago, then the dummy is applied for 1 month ago but
  not 5 months ago).
\end{itemize}

Using this approach, our panel data looks like this:

\begin{verbatim}
# A tibble: 6 x 6
  oracle_number group            year_month absent  days months_since_men~
  <chr>         <chr>            <chr>       <dbl> <int> <fct>            
1 X045753       Temporary field~ 2015_10       0      13 01               
2 X040254       Temporary field~ 2016_9        0      30 01               
3 X040343       Temporary field~ 2014_6        0      30 03               
4 PF02071       Temporary not f~ 2015_3       80.6    31 06               
5 PA00795       Permanent field~ 2013_3       67.7    31 02               
6 PF00699       Permanent not f~ 2016_3        0      31 06               
\end{verbatim}

The model formula was described in Menno's email.

\includegraphics{images/menno}{[}width=3cm{]}

Which translates into R as:

\begin{verbatim}
felm(formula = absent ~ months_since_menno | oracle_number + 
    year_month | 0 | 0, data = these_data)
\end{verbatim}

Note that the \texttt{months\_since\_menno} variable has the following
categories, and is treated as a ``wide'' dummy by the software in
modeling:

\begin{verbatim}
[1] "Before" "01"     "02"     "03"     "04"     "05"     "06"    
\end{verbatim}

The results of this model look like this:

\begin{longtable}[t]{ll}
\caption{\label{tab:unnamed-chunk-8}Simple models with year-month fixed effects}\\
\toprule
Term & Estimate\\
\midrule
\endfirsthead
\caption[]{Simple models with year-month fixed effects \textit{(continued)}}\\
\toprule
Term & Estimate\\
\midrule
\endhead
\
\endfoot
\bottomrule
\endlastfoot
\addlinespace[1.5em]
\multicolumn{2}{l}{\textbf{Permanent field worker}}\\
\hspace{1em}IRS status=\_menno01 & 1.425 (P=0.446)\\
\hspace{1em}IRS status=\_menno02 & 3.603 (P=0.066)\\
\hspace{1em}IRS status=\_menno03 & 2.049 (P=0.297)\\
\hspace{1em}IRS status=\_menno04 & 1.357 (P=0.491)\\
\hspace{1em}IRS status=\_menno05 & 1.351 (P=0.498)\\
\hspace{1em}IRS status=\_menno06 & 3.358 (P=0.106)\\
\addlinespace[1.5em]
\multicolumn{2}{l}{\textbf{Permanent not field worker}}\\
\hspace{1em}IRS status=\_menno01 & -0.022 (P=0.982)\\
\hspace{1em}IRS status=\_menno02 & 0.255 (P=0.805)\\
\hspace{1em}IRS status=\_menno03 & 1.803 (P=0.085)\\
\hspace{1em}IRS status=\_menno04 & -0.87 (P=0.406)\\
\hspace{1em}IRS status=\_menno05 & 0.144 (P=0.894)\\
\hspace{1em}IRS status=\_menno06 & 0.073 (P=0.947)\\
\addlinespace[1.5em]
\multicolumn{2}{l}{\textbf{Temporary field worker}}\\
\hspace{1em}IRS status=\_menno01 & -0.333 (P=0.261)\\
\hspace{1em}IRS status=\_menno02 & -0.058 (P=0.847)\\
\hspace{1em}IRS status=\_menno03 & -0.378 (P=0.215)\\
\hspace{1em}IRS status=\_menno04 & 0.01 (P=0.973)\\
\hspace{1em}IRS status=\_menno05 & 0.075 (P=0.797)\\
\hspace{1em}IRS status=\_menno06 & 0.41 (P=0.17)\\
\addlinespace[1.5em]
\multicolumn{2}{l}{\textbf{Temporary not field worker}}\\
\hspace{1em}IRS status=\_menno01 & 0.08 (P=0.979)\\
\hspace{1em}IRS status=\_menno02 & 2.989 (P=0.35)\\
\hspace{1em}IRS status=\_menno03 & -4.094 (P=0.167)\\
\hspace{1em}IRS status=\_menno04 & -2.276 (P=0.449)\\
\hspace{1em}IRS status=\_menno05 & -2.259 (P=0.464)\\
\hspace{1em}IRS status=\_menno06 & 2.409 (P=0.427)\\*
\end{longtable}

In the above, it is apparent that the effects of IRS appear to be either
(a) in the opposite direction as expected or (b) non-signficant. Here is
a visualization of IRS' apparent effect (relative to the ``Before''
period, which is the reference class). Were IRS effective, we would
expect the black line to be below 0 (the red line).

\begin{center}\includegraphics{menno_models_files/figure-latex/unnamed-chunk-9-1} \end{center}

What is going on here? Could it be that IRS has a positive (ie,
increases) impact on absenteeism? Let's look at the raw data, rather
than the coefficients. The below chart shows absenteeism (as a
percentage of the pre-IRS absenteeism average) for workers, broken down
by season (row) and worker type (column), with point size reflecting the
number of observations.

\begin{center}\includegraphics{menno_models_files/figure-latex/unnamed-chunk-10-1} \end{center}

In the above chart, two things stand out:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Variability is greatest among temporary workers.
\item
  There appears to be a reduction in absenteeism post IRS for permanent
  workers \emph{during the malaria season}, but not during the
  non-malaria season
\end{enumerate}

So, perhaps controlling for year-month through fixed effects isn't
enough, since the effect of IRS \emph{interacts} with the season\ldots{}

\subsection{Model B (with seasonality)}\label{model-b-with-seasonality}

Using this approach, our panel data looks like this:

\begin{verbatim}
# A tibble: 6 x 7
  oracle_number group      year_month absent  days months_since_me~ season
  <chr>         <chr>      <chr>       <dbl> <int> <fct>            <fct> 
1 X037697       Temporary~ 2016_7       0       31 05               low   
2 PF01761       Permanent~ 2015_12      3.23    31 04               high  
3 X040726       Temporary~ 2015_5       3.23    31 02               high  
4 PF00774       Permanent~ 2014_6       3.33    30 04               low   
5 X043249       Temporary~ 2016_4       0       30 06               high  
6 X037700       Permanent~ 2013_2       0       27 02               high  
\end{verbatim}

Our new model specification (in R) is:

\begin{verbatim}
felm(formula = absent ~ months_since_menno * season | oracle_number + 
    year_month | 0 | 0, data = these_data)
\end{verbatim}

The results of this model look like this:

\begin{longtable}[t]{ll}
\caption{\label{tab:unnamed-chunk-13}Simple models with year-month fixed effects and seasonality-IRS interaction term}\\
\toprule
Term & Estimate\\
\midrule
\endfirsthead
\caption[]{Simple models with year-month fixed effects and seasonality-IRS interaction term \textit{(continued)}}\\
\toprule
Term & Estimate\\
\midrule
\endhead
\
\endfoot
\bottomrule
\endlastfoot
\addlinespace[1.5em]
\multicolumn{2}{l}{\textbf{Permanent field worker}}\\
\hspace{1em}IRS status=\_menno01 & 175.053 (P=0.532)\\
\hspace{1em}IRS status=\_menno02 & 457.807 (P=0.115)\\
\hspace{1em}IRS status=\_menno03 & 505.425 (P=0.061)\\
\hspace{1em}IRS status=\_menno04 & 233.373 (P=0.357)\\
\hspace{1em}IRS status=\_menno05 & 227.399 (P=0.372)\\
\hspace{1em}IRS status=\_menno06 & 425.265 (P=0.121)\\
\hspace{1em}Malaria season & 887.601 (P=0.703)\\
\hspace{1em}IRS status=\_menno01:Malaria season & -58.338 (P=0.872)\\
\hspace{1em}IRS status=\_menno02:Malaria season & -174.777 (P=0.642)\\
\hspace{1em}IRS status=\_menno03:Malaria season & -603.926 (P=0.105)\\
\hspace{1em}IRS status=\_menno04:Malaria season & -220.413 (P=0.56)\\
\hspace{1em}IRS status=\_menno05:Malaria season & -213.141 (P=0.582)\\
\hspace{1em}IRS status=\_menno06:Malaria season & -191.654 (P=0.633)\\
\addlinespace[1.5em]
\multicolumn{2}{l}{\textbf{Permanent not field worker}}\\
\hspace{1em}IRS status=\_menno01 & 39.635 (P=0.788)\\
\hspace{1em}IRS status=\_menno02 & 14.833 (P=0.919)\\
\hspace{1em}IRS status=\_menno03 & 114.878 (P=0.439)\\
\hspace{1em}IRS status=\_menno04 & 55.896 (P=0.694)\\
\hspace{1em}IRS status=\_menno05 & 93.301 (P=0.502)\\
\hspace{1em}IRS status=\_menno06 & 86.299 (P=0.556)\\
\hspace{1em}Malaria season & NaN (PNA)\\
\hspace{1em}IRS status=\_menno01:Malaria season & -72.876 (P=0.707)\\
\hspace{1em}IRS status=\_menno02:Malaria season & 23.995 (P=0.903)\\
\hspace{1em}IRS status=\_menno03:Malaria season & 127.34 (P=0.523)\\
\hspace{1em}IRS status=\_menno04:Malaria season & -296.411 (P=0.137)\\
\hspace{1em}IRS status=\_menno05:Malaria season & -185.639 (P=0.372)\\
\hspace{1em}IRS status=\_menno06:Malaria season & -171.932 (P=0.416)\\
\addlinespace[1.5em]
\multicolumn{2}{l}{\textbf{Temporary field worker}}\\
\hspace{1em}IRS status=\_menno01 & -35.174 (P=0.339)\\
\hspace{1em}IRS status=\_menno02 & -6.152 (P=0.869)\\
\hspace{1em}IRS status=\_menno03 & -12.472 (P=0.732)\\
\hspace{1em}IRS status=\_menno04 & 18.095 (P=0.611)\\
\hspace{1em}IRS status=\_menno05 & 33.497 (P=0.346)\\
\hspace{1em}IRS status=\_menno06 & 9.614 (P=0.816)\\
\hspace{1em}Malaria season & 44.327 (P=0.309)\\
\hspace{1em}IRS status=\_menno01:Malaria season & 8.557 (P=0.881)\\
\hspace{1em}IRS status=\_menno02:Malaria season & 3.8 (P=0.948)\\
\hspace{1em}IRS status=\_menno03:Malaria season & -75.835 (P=0.21)\\
\hspace{1em}IRS status=\_menno04:Malaria season & -51.282 (P=0.397)\\
\hspace{1em}IRS status=\_menno05:Malaria season & -70.353 (P=0.217)\\
\hspace{1em}IRS status=\_menno06:Malaria season & 63.717 (P=0.259)\\
\addlinespace[1.5em]
\multicolumn{2}{l}{\textbf{Temporary not field worker}}\\
\hspace{1em}IRS status=\_menno01 & 414.405 (P=0.319)\\
\hspace{1em}IRS status=\_menno02 & 90.379 (P=0.825)\\
\hspace{1em}IRS status=\_menno03 & 353.397 (P=0.441)\\
\hspace{1em}IRS status=\_menno04 & -565.881 (P=0.232)\\
\hspace{1em}IRS status=\_menno05 & -355.43 (P=0.388)\\
\hspace{1em}IRS status=\_menno06 & 45.199 (P=0.895)\\
\hspace{1em}Malaria season & -459.147 (P=0.519)\\
\hspace{1em}IRS status=\_menno01:Malaria season & -724.867 (P=0.206)\\
\hspace{1em}IRS status=\_menno02:Malaria season & 661.492 (P=0.283)\\
\hspace{1em}IRS status=\_menno03:Malaria season & -1223.913 (P=0.031)\\
\hspace{1em}IRS status=\_menno04:Malaria season & 529.98 (P=0.364)\\
\hspace{1em}IRS status=\_menno05:Malaria season & 324.283 (P=0.578)\\
\hspace{1em}IRS status=\_menno06:Malaria season & 817.731 (P=0.227)\\*
\end{longtable}

The below shows model coefficients:

\begin{center}\includegraphics{menno_models_files/figure-latex/unnamed-chunk-14-1} \end{center}

\subsection{Model C (with seasonality and herd
score)}\label{model-c-with-seasonality-and-herd-score}

We now introduce herd immunity to the model. This is introduced as an
additive effect; perhaps it should be interactive with IRS? Using this
approach, our panel data looks like this:

\begin{verbatim}
# A tibble: 6 x 8
  oracle_number group       year_month absent  days  herd months_since_me~
  <chr>         <chr>       <chr>       <dbl> <int> <dbl> <fct>           
1 PF02492       Permanent ~ 2013_3       9.68    31 136   01              
2 X034289       Temporary ~ 2016_9       0       30 276   02              
3 PF00699       Permanent ~ 2016_9       0       30 282   03              
4 X044684       Temporary ~ 2014_9       0       30 248   01              
5 X045133       Temporary ~ 2015_8       6.45    31 156   06              
6 X041009       Temporary ~ 2015_6       0       30  97.1 02              
# ... with 1 more variable: season <fct>
\end{verbatim}

Our new model specification (in R) is:

\begin{verbatim}
felm(formula = absent ~ months_since_menno * season | oracle_number + 
    year_month | 0 | 0, data = these_data)
\end{verbatim}

The results of this model look like this:

\begin{longtable}[t]{ll}
\caption{\label{tab:unnamed-chunk-17}Simple models with year-month fixed effects, seasonality-IRS interaction term, and herd score}\\
\toprule
Term & Estimate\\
\midrule
\endfirsthead
\caption[]{Simple models with year-month fixed effects, seasonality-IRS interaction term, and herd score \textit{(continued)}}\\
\toprule
Term & Estimate\\
\midrule
\endhead
\
\endfoot
\bottomrule
\endlastfoot
\addlinespace[1.5em]
\multicolumn{2}{l}{\textbf{Permanent field worker}}\\
\hspace{1em}IRS status=\_menno01 & 175.053 (P=0.532)\\
\hspace{1em}IRS status=\_menno02 & 457.807 (P=0.115)\\
\hspace{1em}IRS status=\_menno03 & 505.425 (P=0.061)\\
\hspace{1em}IRS status=\_menno04 & 233.373 (P=0.357)\\
\hspace{1em}IRS status=\_menno05 & 227.399 (P=0.372)\\
\hspace{1em}IRS status=\_menno06 & 425.265 (P=0.121)\\
\hspace{1em}Malaria season & 887.601 (P=0.703)\\
\hspace{1em}IRS status=\_menno01:Malaria season & -58.338 (P=0.872)\\
\hspace{1em}IRS status=\_menno02:Malaria season & -174.777 (P=0.642)\\
\hspace{1em}IRS status=\_menno03:Malaria season & -603.926 (P=0.105)\\
\hspace{1em}IRS status=\_menno04:Malaria season & -220.413 (P=0.56)\\
\hspace{1em}IRS status=\_menno05:Malaria season & -213.141 (P=0.582)\\
\hspace{1em}IRS status=\_menno06:Malaria season & -191.654 (P=0.633)\\
\addlinespace[1.5em]
\multicolumn{2}{l}{\textbf{Permanent not field worker}}\\
\hspace{1em}IRS status=\_menno01 & 39.635 (P=0.788)\\
\hspace{1em}IRS status=\_menno02 & 14.833 (P=0.919)\\
\hspace{1em}IRS status=\_menno03 & 114.878 (P=0.439)\\
\hspace{1em}IRS status=\_menno04 & 55.896 (P=0.694)\\
\hspace{1em}IRS status=\_menno05 & 93.301 (P=0.502)\\
\hspace{1em}IRS status=\_menno06 & 86.299 (P=0.556)\\
\hspace{1em}Malaria season & NaN (PNA)\\
\hspace{1em}IRS status=\_menno01:Malaria season & -72.876 (P=0.707)\\
\hspace{1em}IRS status=\_menno02:Malaria season & 23.995 (P=0.903)\\
\hspace{1em}IRS status=\_menno03:Malaria season & 127.34 (P=0.523)\\
\hspace{1em}IRS status=\_menno04:Malaria season & -296.411 (P=0.137)\\
\hspace{1em}IRS status=\_menno05:Malaria season & -185.639 (P=0.372)\\
\hspace{1em}IRS status=\_menno06:Malaria season & -171.932 (P=0.416)\\
\addlinespace[1.5em]
\multicolumn{2}{l}{\textbf{Temporary field worker}}\\
\hspace{1em}IRS status=\_menno01 & -35.174 (P=0.339)\\
\hspace{1em}IRS status=\_menno02 & -6.152 (P=0.869)\\
\hspace{1em}IRS status=\_menno03 & -12.472 (P=0.732)\\
\hspace{1em}IRS status=\_menno04 & 18.095 (P=0.611)\\
\hspace{1em}IRS status=\_menno05 & 33.497 (P=0.346)\\
\hspace{1em}IRS status=\_menno06 & 9.614 (P=0.816)\\
\hspace{1em}Malaria season & 44.327 (P=0.309)\\
\hspace{1em}IRS status=\_menno01:Malaria season & 8.557 (P=0.881)\\
\hspace{1em}IRS status=\_menno02:Malaria season & 3.8 (P=0.948)\\
\hspace{1em}IRS status=\_menno03:Malaria season & -75.835 (P=0.21)\\
\hspace{1em}IRS status=\_menno04:Malaria season & -51.282 (P=0.397)\\
\hspace{1em}IRS status=\_menno05:Malaria season & -70.353 (P=0.217)\\
\hspace{1em}IRS status=\_menno06:Malaria season & 63.717 (P=0.259)\\
\addlinespace[1.5em]
\multicolumn{2}{l}{\textbf{Temporary not field worker}}\\
\hspace{1em}IRS status=\_menno01 & 414.405 (P=0.319)\\
\hspace{1em}IRS status=\_menno02 & 90.379 (P=0.825)\\
\hspace{1em}IRS status=\_menno03 & 353.397 (P=0.441)\\
\hspace{1em}IRS status=\_menno04 & -565.881 (P=0.232)\\
\hspace{1em}IRS status=\_menno05 & -355.43 (P=0.388)\\
\hspace{1em}IRS status=\_menno06 & 45.199 (P=0.895)\\
\hspace{1em}Malaria season & -459.147 (P=0.519)\\
\hspace{1em}IRS status=\_menno01:Malaria season & -724.867 (P=0.206)\\
\hspace{1em}IRS status=\_menno02:Malaria season & 661.492 (P=0.283)\\
\hspace{1em}IRS status=\_menno03:Malaria season & -1223.913 (P=0.031)\\
\hspace{1em}IRS status=\_menno04:Malaria season & 529.98 (P=0.364)\\
\hspace{1em}IRS status=\_menno05:Malaria season & 324.283 (P=0.578)\\
\hspace{1em}IRS status=\_menno06:Malaria season & 817.731 (P=0.227)\\*
\end{longtable}

>>>>

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Concept note: Maragra model approach - September 2018},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

\setlength{\droptitle}{-2em}

  \title{Concept note: Maragra model approach - September 2018}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \subtitle{Pradhan, Sicuri, Brew}
  \author{}
    \preauthor{}\postauthor{}
    \date{}
    \predate{}\postdate{}

@@

Maragra models

@@

This document contains methods and basic results for a few different modeling approaches. It was produced in September 2018 in Amsterdam. The purpose of this document is to give an overview of the modeling approached employed so far, as well as provide a summary of results.

@@

\noindent\fbox{%
    \parbox{\textwidth}{%
        \subsection*{Main points}
        \begin{itemize}
          \item We have combined individual and indirect protection into one index
          \item We have simplified our model approach
          \item Results are consistent/coherent
        \end{itemize}
        \vspace{2mm}
    }%
}

@@

\section{The previous approach}\label{the-previous-approach}

In the draft of the paper
(\href{https://docs.google.com/document/d/1bUWRBCgVcgjSPHchIQxiTG8Vwv5hV1GLU4Tlu386sWA/edit#}{HERE}
), the model formula is:

@@

\begin{verbatim}
felm(formula = absent ~ season * months_since + rainy_day + herd | 
    oracle_number + malaria_year | 0 | 0, data = these_daily_data)
\end{verbatim}

@@

\section{The new approach}\label{the-new-approach}

The new approach, discussed September 26 2018, is characterized by the
following:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  We assume that the sum of ``spraying'' and rainfall is what determines
  the quantity of malaria-infectious mosquitoes at any given time.
\item
  We assume that rainfall's effect is best captured by lagging 4-13
  weeks prior to the date of observation (Wu et al. 2017).
\item
  We assume that the relationship between this sum takes on an
  exponential shape.
\item
  We model separately for worker types (permanent vs.~temporary and
  factory vs.~field).
\item
  We assume that other factors (age, sex, etc.) not captured by the
  above do not interact with absenteeism or the effect of IRS.
\item
  We assume that the protection afforded by IRS is a function of
  distance to that IRS, and are indifferent to (a) the relative coverage
  of IRS in your area (the absolute quantity of chemical being the only
  thing that matters) and (b) whether the IRS is at your house or
  elsewhere (the distance to the IRS being the only thing that matters)
\end{enumerate}

\subsection{Specification}\label{specification}

\[
\hat{Y_{it}} = \hat{\beta}_{0} +  \hat{\beta}_{1}\text{exp(protection)}_{it}  + \hat{\beta}_{2}\text{Rain}_{it} +  \alpha_i + \upsilon_{it}
\]

The above notation shows \texttt{absent} (binary 1,0) as a function of
the exponentiation of the \texttt{protection} variable (the summation of
individual plus indirectly conferred protection) and the \texttt{rain}
variable (the summed lag of rainfall from 4-13 weeks). \(\alpha_i\)
represents the time invariant worker fixed effects, \(\upsilon\) is the
error term.

\subsection{Code}\label{code}

In R, our new model looks like this

\begin{verbatim}
felm(formula = log(absent + 1) ~ protection + precipitation_lag1_3 | 
    oracle_number | 0 | 0, data = monthly)
\end{verbatim}

Where \texttt{protection} is the sum of all nearby IRS, weighted by the
inverse of the distance from the residence of the person (including the
person's own residence, set at a distance of 10 meters).

\subsection{Data structure}\label{data-structure}

\begin{verbatim}
# A tibble: 6 x 29
  oracle_number year_month group precipitation absences longitude_aura
  <chr>         <chr>      <chr>         <dbl>    <int>          <dbl>
1 CA14693       2013-01-01 Temp~          10.7        0           32.7
2 CA15945       2013-01-01 Temp~          10.7        0           NA  
3 CA16471       2013-01-01 Perm~          10.7        2           32.8
4 CA16473       2013-01-01 Perm~          10.7       11           32.8
5 CA16476       2013-01-01 Perm~          10.7       11           NA  
6 CA16477       2013-01-01 Perm~          10.7        0           32.8
# ... with 23 more variables: latitude_aura <dbl>, ever_sprayed <lgl>,
#   eligibles <int>, herd <dbl>, protection <dbl>, irs <dbl>, sex <chr>,
#   permanent_or_temporary <chr>, high_season <dbl>, absent <dbl>,
#   precipitation_lag0 <dbl>, precipitation_lag1 <dbl>,
#   precipitation_lag2 <dbl>, precipitation_lag3 <dbl>,
#   precipitation_lag4 <dbl>, precipitation_lag2_3 <dbl>,
#   precipitation_lag1_2 <dbl>, precipitation_lag0_2 <dbl>,
#   precipitation_lag0_3 <dbl>, precipitation_lag1_3 <dbl>,
#   precipitation_lag1_4 <dbl>, herd_ideal <dbl>, protection_ideal <dbl>
\end{verbatim}

\subsection{Data summary}\label{data-summary}

\begin{center}\includegraphics{sep2018_files/figure-latex/unnamed-chunk-7-1} \end{center}

\begin{center}\includegraphics{sep2018_files/figure-latex/unnamed-chunk-8-1} \end{center}

\begin{center}\includegraphics{sep2018_files/figure-latex/unnamed-chunk-9-1} \end{center}

\subsection{Results}\label{results}

The below table shows the results of the model devised thus far.

\begin{longtable}[t]{ll}
\caption{\label{tab:unnamed-chunk-10}Model results}\\
\toprule
Term & Estimate\\
\midrule
protection & -0.192 (P=0.021)\\
Precipitation (mm) \_lag1\_3 & 0.04 (P=0.027)\\
\bottomrule
\end{longtable}

\subsection{Intepretation}\label{intepretation}

The below charts show predicted absenteeism rates at different rain and
protection levels.

\subsubsection{Precipitation's effect}\label{precipitations-effect}

\begin{center}\includegraphics{sep2018_files/figure-latex/unnamed-chunk-12-1} \end{center}

\subsubsection{IRS protection's effect}\label{irs-protections-effect}

\begin{center}\includegraphics{sep2018_files/figure-latex/unnamed-chunk-13-1} \end{center}

\section*{Bibliography}\label{bibliography}
\addcontentsline{toc}{section}{Bibliography}

\hypertarget{refs}{}
\hypertarget{ref-Wu_2017}{}
Wu, Yunyun, Zhijiao Qiao, Nan Wang, Hongjie Yu, Zijian Feng, Xiaosong
Li, and Xing Zhao. 2017. ``Describing Interaction Effect Between Lagged
Rainfalls on Malaria: An Epidemiological Study in South--West China.''
\emph{Malaria Journal} 16 (1). Springer Nature.
doi:\href{https://doi.org/10.1186/s12936-017-1706-2}{10.1186/s12936-017-1706-2}.
